The Hot Hand Fallacy in Call of Duty Search and Destroy

### 2022-08-30 paper/paper.tex
final? with gt

\begin{abstract}
Our research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, we find evidence
supporting the hypothesis that round win probability can be modeled as a
constant across rounds in the series, although not at the naive 50\%.
Second, we examine the proportion of round wins immediately following a
win streak. Given streak length, series length, and series winner, we
find no evidence that post-streak win proportion is significantly
different than expected, suggesting that the perception of momentum at
the team level is deceiving. Wald-Wolfowitz run tests also fail to
provide evidence for the hot hand phenomenon.
\end{abstract}

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. The most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
in which one team tries to destroy one of two designated bomb sites on
the map.\footnote{SnD bears resemblance to ``Bomb Defusal'' in
  Counter-Strike and ``Plant/Defuse'' in Valorant, two other FPS games
  played in more popular professional leagues.}

In professional CoD SnD, the two teams\footnote{In the 2020 season,
  teams played with five players each; in the 2021 and 2022 seasons,
  teams played with four players each.} take turns playing offense and
defense every round. The first to six round wins (best-of-11 round
format) is declared the series winner.\footnote{A short list of rules
  that govern timing follows.} A round can end in one of four ways:

\begin{itemize}
\tightlist
\item
  Each round has a 90 second time limit, not counting the potential
  extension granted if a bomb is planted
\item
  A bomb plant takes five seconds. The timer resets if the player stops
  planting site prior to completing it.
\item
  A bomb defuse takes seven seconds. The timer resets if the player
  ``drops'' the bomb.
\item
  Once the bomb is planted, the round timer is set to 45 seconds.
\end{itemize}

@@

We adopt the terminology ``series'' to refer to a single best-of-11
matchup, so as to mirror the terminology of playoff series in
professional leagues like the National Basketball Association (NBA),
National Hockey League (NHL), and Major League Baseball (MLB). A
``game'' or a ``match'' in such leagues is analogous to a ``round'' of
CoD SnD for the purposes of this paper. Note that SnD is not the only
game mode played in competitive CoD.\footnote{Hardpoint has been played
  as the first and fourth game modes in a matchup. The third game mode
  was Domination in the 2020 season, and Control in the 2021 and 2022
  seasons.} Teams play in a head-to-head, best-of-five format, where SnD
is always played as the second and fifth game modes. The best-of-five
matchup could also be called a ``series'', but since we analyze only the
SnD games, we refer to the SnD games as series.

@@

The data set consists of all SnD matches played in major tournaments and
qualifiers during the past three years\footnote{Since the release of
  ``Call of Duty: Modern Warfare'' in fall 2019, professional CoD has
  been orchestrated by the 12-franchise CoD League (CDL). The CDL has
  completed three year-long ``seasons'' as of August 2022.}, totaling
7,792 rounds across 852 series.\footnote{288 of the series occur in
  major tournaments, which are considered to be more competitive than
  the qualifiers since they are played in person (COVID-permitting),
  eliminating randomness due to online latency and ping. The qualifiers
  are played online.} Data has been collected in spreadsheets by
community member ``IOUTurtle''.\footnote{Raw data:
  \url{https://linktr.ee/CDLArchive}. Cleaned data:
  \url{https://github.com/\%7Bauthor\%7D/\%7Brepo\%7D/blob/master/data/\%7Bfile\%7D.csv}.}

CoD is fairly unique compared to other esports in that it runs on an
annual lifecycle, with releases coming in the late fall. A new game
under the same brand---``Call of Duty''---is published every year by a
rotating set of developers. Each new game bears resemblance to past
ones, often introducing relatively small variations (``improvements'')
to graphics, game modes, and other facets of gameplay. During the CDL
era, the games released have been ``Modern Warfare'' (2020 season),
``Cold War'' (2021 season) and ``Vanguard'' (2022 season).

Sweeps (6-0 series) make up 4.7\% of all series, while 6-5 series make
up 23.7\% of series. The observed offensive round win
percentage\footnote{We use the terms ``percentage'', ``probability'',
  and ``proportion'' synonymously throughout this paper when discussing
  round win rates. For the most part, we restrict ``percentage'' usage
  to observed values and ``probability'' usage to expected values, and
  we use ``proportion'' to refer to either observed or expected values.}
across all rounds is \(\bar{\tau}\) = 47.8\%.\footnote{Offensive round
  win percentage has been nearly constant across the three games during
  the CDL era: 47.2\% in Modern Warfare; 47.9\% in Cold War; and 48.1\%
  in Vanguard} Table \ref{tbl:o-win-prop-by-series-state} shows
offensive round win percentages by series ``state'', i.e.~the number of
round wins by each team prior to an upcoming round. Offensive round win
percentage is not quite constant, although never veers more than 10\%
from \(\bar{\tau}\).

@@

The professional competition format balances the frequency with which
teams play offense or defense to start a series, to the extent possible.
Thus, win percentages in certain states are subjected to minimal
selection bias, i.e.~bias due to better teams playing defense more
frequently.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense in the SnD series, played
  as the second or fifth game mode of the best-of-five matchup. Their
  opponent is assigned to start on offense in the SnD slot not chosen.
  Anecdotally, the better team tends to choose to play defense in the
  first round of the SnD series played as the second game mode in the
  best-of-five matchup, although this choice is not consistent across
  teams, or even with the same team over time.}

@@

This format has minimized the range of rates at which teams start series
playing offense. The highest rate at which a team has started their
series playing defense across all tournament and qualifier SnD series in
a single season is 66\% (31 of 47). The lowest is 38\% (16 of 42).

@@

Other research has borrowed techniques from the field of quality
control, such as identifying unlikely sequences of events with the
Wald-Wolfowitz runs test (Peel and Clauset 2015, Steeger et al.~2021).
Peel and Clauset found no evidence for unlikely sequences of scoring
events in the NHL, College Football, and National Football League,
although did in the NBA. As a check on their entropy approach to
momentum identification, Steeger et al.~found several NHL teams with
sequences of wins in the 2018-2019 regular season that violated the
Wald-Wolfowitz null hypothesis.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  team in a round, given how prior rounds played out. CoD has no such
  equivalent, except for perhaps ``score streaks'', which infrequently
  occur.
\item
  Teams play either offense or defense for many consecutive rounds.
  (Counter-Strike has a 15-15-1 format and Valorant has a 12-12-1
  format.) On the other hand, teams in CoD SnD rotate attacking and
  defending roles every round, minimizing the effect of map asymmetry on
  streaks of wins.\footnote{Professional CoD has a set of maps (at least
    three, depending on the season) on which teams choose to play SnD.
    Most are asymmetric, introducing slight advantages for the offense
    or defense that do not exist on other maps. Counter-Strike and
    Valorant also have their own sets of maps for their analogous SnD
    modes.}
\item
  Both Counter-Strike and Valorant have overtime rules---a team must win
  by two rounds---which can make end-of-series sequences difficult to
  model properly. CoD SnD does not have overtime rules.
\end{enumerate}

@@

As implied by MS's Theorem 1, we should expect the proportion of round
wins immediately following a streak of \(k\) rounds wins for a series
ending in \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly less
than the observed proportion of round wins given the series length, but
not given knowledge of the potential existence of a streak, \(P^{+|r}\),
i.e.~the analogue of the expected proportion in Equation \ref{eq:pwr}
for the observed data. However, our context is fundamentally different
from that of MS, who evaluate data in controlled settings, unaffected by
opposition.\footnote{GVT also perform statistical tests on shots from
  players in live games, i.e.~``observational'' data, but they note that
  their findings are likely affected by player shot selection influenced
  by the opposing team's defensive strategy. This is similar to our
  setting.}

@@

Let us consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
ending in \(r\) rounds: the ``notional'' proportion

\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
1 & \text{team wins series}, r = m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}

where \(r \in R\) and \(m\) from Equation \ref{eq:m}.

@@

We can further decompose Equation \ref{eq:pwkr} by the round \(i\)
(where \(k < i \leq r\)) in which the streak of length \(k\) carries
into.

\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, i \neq r; \\
1 & \text{team wins series}, i = r; \\
\frac{s - m - k}{r - k} & \text{team loses series}, i \neq r, r > m + k; \\
0 & \text{team loses series}, i = r.
\end{cases}
\end{equation}

Again, we can apply a binomial test to evaluate the hypothesis that the
expected proportion, \(\hat{P}^{+|k,r,i}_0\), is equal to the observed
proportion, \(P^{+|k,r,i}\).

@@

We can attempt to detect the hot hand phenomenon more broadly by
focusing on non-random sequences identified by the Wald-Wolfowitz runs
test. Note that the probability of success and failure need not be
equal; thus, we can incorporate our findings regarding constant round
win probability, specifying that the probability of success is
\(\phi_2 = 0.575\).

@@

We can subset the observed series sequences, \(\zeta_j\), to those that
violate the null hypothesis for the runs test and perform a test of
equal proportions, where the null hypothesis is that the observed
proportion of a sequence relative to all possible sequences,
\(P^{\zeta}_j\) ( \(\frac{N^\zeta_j}{\sum_j N^\zeta }\)), is equal to
the expected proportion of the sequence relative to all possible
sequences, \(\hat{P}^\zeta_j\). The test statistic is

@@

where \(N^\zeta_j\) and \(\hat{N}^\zeta_j\) respectively represent the
observed and expected number of occurrences of sequence \(\zeta_j\), and
where
\(P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})\).
The null hypothesis is that the test statistic is normally distributed.
If we can reject the null hypothesis for such sequences, we can build an
argument in support of the hot hand effect.\footnote{As with the
  binomial tests for the notional and MS post-streak proportions, we
  adjust the \(p\)-values with the BY correction.}

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion
of series ending in \(r\) rounds given \(\phi_0 = 0.5\),
\(\hat{\Phi}_0(r)\), as well as the observed proportions, \(\Phi(r)\).

@@

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum
(\protect\hyperlink{ref-carlson2007}{Carlson and Shu 2007}), we focus on
streaks of three round wins. Table \ref{tbl:pwkr} compares the observed
round win proportion, \(P^{+|k=3,r}\), with the notional proportion,
\(\hat{P}^{+|k=3,r}\), following streaks of \(k=3\) round wins. The MS
proportion, \(\hat{P}^{+|k=3,r}_{MS}\), is also shown as a reference.

@@

All binomial test \(p\)-values for both \(\hat{P}^{+|k=3,r}_0\) and
\(\hat{P}^{+|k=3,r}_{MS}\) are found to be insignificant at the
\(\alpha = 0.05\) threshold, implying that we cannot reject the null
hypothesis that the notional and MS post-streak round win proportions
are different than the observed proportion for streaks of
three.\footnote{\(p\)-values are not shown for readability purposes.}
When performing the same tests for streaks of two, four, and five, there
is no case in which we can reject the null hypothesis for
\(\hat{P}^{+|k,r}_0\). The null hypothesis can only be rejected in some
extreme cases for \(\hat{P}^{+|k,r}_{MS}\).\footnote{These cases are:
  (1) \(k = 4\), \(r = 7\) (\(N^{k,r} = 108\)); (2) \(k = 4\),
  \(r = 10\) (\(N^{k,r} = 45\)); (3) \(k = 5, r = 7\)
  (\(N^{k,r} = 37\)); (4) \(k = 5, r = 8\) (\(N^{k,r} = 4\)). In each
  case, the team won the series.}

Table \ref{tbl:pw3ri} shows the observed and notional proportions of
round wins immediately following streaks of three round wins when
explicitly accounting for the round index, \(i\).\footnote{We omit the
  MS proportions \(\hat{P}^{+|k=3,r,i}_{MS}\) since sample sizes are
  small and results may or may not be valid due to the questionable
  i.i.d assumption.}

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r,i}$, and proportion of round wins, $P^{+|k=3,r,i}$, among $N^{k=3,r,i}$ chances, given the index of the round, $i$, immediately following the streak. Additionally, the notional proportion, $\hat{P}^{+|k=3,r,i}_0$. Table restricted to $r \in [10, 11]$, $i < r$, and $N^{k=3,r,i} \geq 10$.}

\centering
\begin{tabular}{rrcrrrr}
\toprule
$r$ & $i$ & \text{Win series?} & $r^{+|k=3,r,i}$ & $N^{k=3,r,i}$ & $P^{+|k=3,r,i}$ & $\hat{P}^{+|k=3,r,i}_0$\\ 
\midrule

10 & 7 & no & 2 & 13 & 15.4\% & 14.3\% \\ 
10 & 7 & yes & 11 & 13 & 84.6\% & 42.9\% \\ 
10 & 8 & no & 1 & 10 & 10.0\% & 14.3\% \\ 
10 & 8 & yes & 9 & 10 & 90.0\% & 42.9\% \\ 
11 & 5 & no & 9 & 13 & 69.2\% & 25.0\% \\ 
11 & 5 & yes & 4 & 13 & 30.8\% & 37.5\% \\ 
11 & 6 & no & 4 & 13 & 30.8\% & 25.0\% \\ 
11 & 6 & yes & 9 & 13 & 69.2\% & 37.5\% \\ 
11 & 7 & no & 3 & 10 & 30.0\% & 25.0\% \\ 
11 & 7 & yes & 7 & 10 & 70.0\% & 37.5\% \\ 
11 & 9 & no & 5 & 15 & 33.3\% & 25.0\% \\ 
11 & 9 & yes & 10 & 15 & 66.7\% & 37.5\% \\ 
11 & 10 & no & 5 & 10 & 50.0\% & 25.0\% \\ 
11 & 10 & yes & 5 & 10 & 50.0\% & 37.5\% \\

@@

In Table \ref{tbl:ww-sequences}, the observed and expected count of
sequences respectively, are shown for sequence for which we can reject
the Wald-Wolfowitz null hypothesis\footnote{Interestingly, we identify
  the same non-random sequences when using either \(\phi = 0.575\) or
  \(\phi = 0.5\).} at a confidence level of \(\alpha = 0.05\), for
\(r \in [6, 7, 8, 9, 10]\)\footnote{Since the normal method is invalid
  for sequences with only successes or failures, we use the exact method
  to evaluate (and reject) the null hypothesis for the Wald Wolfowitz
  runs test for \(r = 6\).}. In addition to the 14 sequences shown,
there are 12 additional sequences for \(r = 11\).

@@

\caption{Sequences ($\zeta_j$) for which we can reject Wald-Wolfowitz null hypothesis for $r \in [6, 7, 8, 9, 10]$, where round wins and losses are denoted with $+$ and $-$ respectively. The observed count, $N^{\zeta}_j$, and proportion, $P^{\zeta}_j$, of all possible sequences, as well as the expected count, $\hat{N}^{\zeta}_j$, and proportion, $\hat{P}^{\zeta}_j$. $\hat{N}^{\zeta}_j$ is scaled to the observed number of series played, hence its non-integer value.}

>>>>

\begin{abstract}
Our research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, we find evidence
supporting the hypothesis that round win probability can be modeled as a
constant across rounds in the series, although not at the naive 50\%.
Second, we examine the proportion of round wins immediately following a
win streak. Given the round index, streak length, series length, and
series winner, we find no evidence that post-streak win proportion is
significantly different than expected, suggesting that the perception of
momentum at the team level is deceiving. Finally, we find that sequences
of round wins that are identified as non-random by the Wald-Wolfowitz
run tests do not occur more or less frequently than expectated, further
failing to provide evidence for the hot hand effect.
\end{abstract}

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. The most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
in which one team tries to destroy one of two designated bomb sites on
the map.\footnote{SnD bears resemblance to ``Bomb Defusal'' in
  Counter-Strike and ``Plant/Defuse'' in Valorant, two other FPS games
  played in popular professional leagues.}

In professional CoD SnD, the two teams\footnote{In the 2020 season,
  teams played with five players each; in the 2021 and 2022 seasons,
  teams played with four players each.} take turns playing offense and
defense every round. The first to six round wins (best-of-11 round
format) is declared the series winner.\footnote{A short list of rules
  that govern timing follows.

  \begin{itemize}
  \tightlist
  \item
    Each round has a 90 second time limit, not counting the potential
    extension granted if a bomb is planted
  \item
    A bomb plant takes five seconds. The timer resets if the player
    stops planting site prior to completing it.
  \item
    A bomb defuse takes seven seconds. The timer resets if the player
    ``drops'' the bomb.
  \item
    Once the bomb is planted, the round timer is set to 45 seconds,
    regardless of the prior round timer.
  \end{itemize}} A round can end in one of four ways:

@@

We adopt the terminology ``series'' to refer to a single best-of-11
matchup, so as to mirror the terminology of playoff series in
professional leagues like the National Basketball Association (NBA),
National Hockey League (NHL), and Major League Baseball (MLB). A
``game'' or a ``match'' in such leagues is analogous to a ``round'' of
CoD SnD for the purposes of this paper.

Note that SnD is not the only game mode played in competitive
CoD.\footnote{Hardpoint has been played as the first and fourth game
  modes in a matchup. The third game mode was Domination in the 2020
  season, and Control in the 2021 and 2022 seasons.} Teams play in a
head-to-head, best-of-five format, where SnD is always played as the
second and fifth game modes. The best-of-five matchup could also be
called a ``series'', but since we analyze only the SnD games, we refer
to the SnD games as series.

@@

The data set consists of all SnD matches played in major tournaments and
qualifiers during the past three years\footnote{Since the release of
  ``Call of Duty: Modern Warfare'' in fall 2019, professional CoD has
  been orchestrated by the 12-franchise CoD League (CDL). The CDL has
  completed three year-long ``seasons'' as of August 2022.

  CoD is fairly unique compared to other esports in that it runs on an
  annual lifecycle, with releases coming in the late fall. A new game
  under the same brand---``Call of Duty''---is published every year by a
  rotating set of developers. Each new game bears resemblance to past
  ones, often introducing relatively small variations (``improvements'')
  to graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been ``Modern Warfare'' (2020 season),
  ``Cold War'' (2021 season) and ``Vanguard'' (2022 season).}, totaling
7,792 rounds across 852 series.\footnote{288 of the series occur in
  major tournaments, which are considered to be more competitive than
  the qualifiers since they are played in person (COVID-permitting),
  eliminating randomness due to online latency and ping. The qualifiers
  are played online.} Data has been collected in spreadsheets by
community member ``IOUTurtle''.\footnote{Raw data:
  \url{https://linktr.ee/CDLArchive}. Cleaned data:
  \url{https://github.com/author/repo/blob/master/data/file.csv}.}

Sweeps (6-0 series) make up 4.7\% of all series, while 6-5 series make
up 23.7\% of series. The observed offensive round win
percentage\footnote{We use the terms ``percentage'', ``probability'',
  and ``proportion'' synonymously throughout this paper when discussing
  round win rates. For the most part, we restrict ``percentage'' usage
  to observed values and ``probability'' usage to expected values, and
  we use ``proportion'' to refer to either observed or expected values.}
across all rounds is \(\bar{\tau}\) = 47.8\%.\footnote{Offensive round
  win percentage has been nearly constant across the three games during
  the CDL era: 47.2\% in the 2020 season; 47.9\% in the 2021 season; and
  48.1\% in the 2022 season.} Table \ref{tbl:o-win-prop-by-series-state}
shows offensive round win percentages by series ``state'', i.e.~the
number of round wins by each team prior to an upcoming round. Offensive
round win percentage is not quite constant, although never veers more
than 10\% from \(\bar{\tau}\).

@@

The professional competition format balances the frequency with which
teams play offense or defense to start a series, to the extent possible.
Thus, win percentages in certain states are subjected to minimal
selection bias, i.e.~bias due to better teams playing defense more
frequently.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense in the SnD series, played
  as the second or fifth game mode of the best-of-five matchup. Their
  opponent is assigned to start on offense in the SnD slot not chosen.
  Anecdotally, the better team tends to choose to play defense in the
  first round of the SnD series played as the second game mode in the
  best-of-five matchup, although this choice is not consistent across
  teams, or even with the same team over time.

@@

  This format has minimized the range of rates at which teams start
  series playing offense. The highest rate at which a team has started
  their series playing defense across all tournament and qualifier SnD
  series in a single season is 66\% (31 of 47 series). The lowest is
  38\% (16 of 42 series).}

@@

Other research has borrowed techniques from the field of quality
control, such as identifying unlikely sequences of events with the
Wald-Wolfowitz runs test. Peel and Clauset
(\protect\hyperlink{ref-peel2015}{2015}) found no evidence for unlikely
sequences of scoring events in the NHL, College Football, and National
Football League, although did in the NBA. As a check on their entropy
approach to momentum identification, Steeger et al.
(\protect\hyperlink{ref-steeger2021}{2021}) found several NHL teams with
sequences of wins in the 2018-2019 regular season that violated the
Wald-Wolfowitz null hypothesis.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  team in a round, given how prior rounds played out. CoD has no such
  equivalent, except for perhaps ``score streaks'', which infrequently
  occur.
\item
  Teams play either offense or defense for many consecutive rounds.
  (Counter-Strike has a 15-15-1 format and Valorant has a 12-12-1
  format.) On the other hand, teams in CoD SnD rotate attacking and
  defending roles every round, minimizing the effect of map asymmetry on
  streaks of wins.\footnote{Professional CoD has a set of maps (at least
    four, depending on the season) on which teams choose to play SnD.
    Most are asymmetric, introducing slight advantages for the offense
    or defense that do not exist on other maps. Counter-Strike and
    Valorant also have their own sets of maps for their analogous SnD
    modes.}
\item
  Both Counter-Strike and Valorant have overtime rules---a team must win
  by two rounds---which can make end-of-series sequences difficult to
  model properly. CoD SnD does not have overtime rules.
\end{enumerate}

@@

As implied by MS's Theorem 1, we should expect the proportion of round
wins immediately following a streak of \(k\) rounds wins for a series
ending in \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly less
than the observed proportion of round wins given the series length,
where knowledge of the potential existence of a streak is not given,
\(P^{+|r}\), i.e.~the analogue of the expected proportion in Equation
\ref{eq:pwr} for the observed data. However, our context is
fundamentally different from that of MS, who evaluate data in controlled
settings, unaffected by opposition.\footnote{GVT also perform
  statistical tests on shots from players in live games,
  i.e.~``observational'' data, but they note that their findings are
  likely affected by player shot selection influenced by the opposing
  team's defensive strategy. This is similar to our setting.}

@@

Let us consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
ending in \(r\) rounds, the ``notional'' proportion. We define the
notional proportion as

\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
1 & \text{team wins series}, r = m; \\
0 & \text{team loses series}, r = m + k; \\
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k,
\end{cases}
\end{equation}

where \(r \in R\) and \(m\) from Equation \ref{eq:m}.\footnote{Note that
  some cases are invalid, such as \(\text{team loses series}\),
  \(r < m + k\). It is impossible for team to go on a streak of \(k\)
  round wins and then lose the series in less than \(m + k\) rounds.}

@@

We can further decompose Equation \ref{eq:pwkr} by the round \(i\) (
\(k < i \leq r\)) in which the streak of length \(k\) carries into:

\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
1 & \text{team wins series}, i = r \text{ or } r = m; \\
0 & \text{team loses series}, i = r \text{ or } r = m + k; \\
\frac{m-k-1}{r-k-1} & i < r-1; \\
\frac{1}{2} & i = r-1.
\end{cases}
\end{equation}

where \(r \in R\) and \(m\) from Equation
\ref{eq:m}.\footnote{The case where \(i = r - 1\) is the round prior to
  the known last round in the series. We know the result of the last
  round since we know which team won the series. The penultimate round
  is essentially treated as a toss-up, hence \(\frac{1}{2}\).}\footnote{As
  with Equation \ref{eq:pwkr}, there are cases that are impossible, such
  as \(\text{team loses series}\), \(r < m + k\).} Again, we can apply a
binomial test to evaluate the hypothesis that the expected proportion,
\(\hat{P}^{+|k,r,i}_0\), is equal to the observed proportion,
\(P^{+|k,r,i}\).

@@

We can attempt to detect the hot hand phenomenon more broadly by
focusing on non-random sequences identified by the Wald-Wolfowitz runs
test. Note that the probability of success and failure need not be
equal; thus, we can incorporate our findings regarding constant round
win probability.

@@

We can subset the observed series sequences, \(\zeta_j\), to those that
violate the null hypothesis for the runs test and perform a test of
equal proportions, where the null hypothesis is that the observed
proportion of a sequence relative to all possible sequences,
\(P^{\zeta}_j\) ( \(N^\zeta_j / \sum_j N^\zeta_j\)), is equal to the
expected proportion of the sequence relative to all possible sequences,
\(\hat{P}^\zeta_j\). The test statistic is

@@

where \(N^\zeta_j\) and \(\hat{N}^\zeta_j\) respectively represent the
observed and expected number of occurrences of a sequence \(\zeta_j\),
and where
\(P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})\).
The null hypothesis is that the test statistic is normally distributed.
If we can reject the null hypothesis for such sequences, we can build an
argument in support of the hot hand effect.\footnote{As with the
  binomial tests for the post-streak proportions, we adjust the
  \(p\)-values with the BY correction.}

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion
of series ending in \(r\) rounds given \(\phi_0\), \(\hat{\Phi}_0(r)\),
as well as the observed proportions, \(\Phi(r)\).

@@

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum
(\protect\hyperlink{ref-carlson2007}{Carlson and Shu 2007}), we focus on
streaks of three round wins. Table \ref{tbl:pwkr} compares the observed
round win proportion, \(P^{+|k=3,r}\), with the notional proportion,
\(\hat{P}^{+|k=3,r}_0\), following streaks of \(k=3\) round wins. The MS
proportion, \(\hat{P}^{+|k=3,r}_{MS}\), is also shown as a reference.

@@

All binomial test \(p\)-values for both \(\hat{P}^{+|k=3,r}_0\) and
\(\hat{P}^{+|k=3,r}_{MS}\) are found to be insignificant at the
\(\alpha = 0.05\) threshold, implying that we cannot reject the null
hypothesis that the notional and MS post-streak round win proportions
are different than the observed proportion for streaks of three. When
performing the same tests for streaks of two, four, and five, there is
no case in which we can reject the null hypothesis for
\(\hat{P}^{+|k,r}_0\). The null hypothesis can only be rejected in some
extreme cases for \(\hat{P}^{+|k,r}_{MS}\).\footnote{These cases are:
  (1) \(k = 4\), \(r = 7\) (\(N^{k,r} = 108\)); (2) \(k = 4\),
  \(r = 10\) (\(N^{k,r} = 45\)); (3) \(k = 5, r = 7\)
  (\(N^{k,r} = 37\)); (4) \(k = 5, r = 8\) (\(N^{k,r} = 4\)). In each
  case, the team won the series.}

Table \ref{tbl:pw3ri} shows the observed and notional proportions of
round wins---\(P^{+|k=3,r,i}\) and \(\hat{P}^{+|k=3,r,i}_0\)
respectively---immediately following streaks of three round wins when
explicitly accounting for the post-streak round index, \(i\).\footnote{We
  omit the MS proportions \(\hat{P}^{+|k=3,r,i}_{MS}\) since sample
  sizes are small and results may or may not be valid due to the
  questionable i.i.d assumption.} While some of the observed and
notional proportions differ strongly, the BY-corrected \(p\)-values
indicate that the null hypothesis cannot be rejected for any combination
of \(i\) and \(r\), including those not shown for brevity.\footnote{There
  are combinations for which we could reject the null hypothesis if we
  did not apply the BY correction, such as for \(r = 10\) and \(i = 9\).}
The same is true for streaks of two, four, and five.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r,i}$, and proportion of round wins, $P^{+|k=3,r,i}$, among $N^{k=3,r,i}$ chances, given the index of the round, $i$, immediately following the streak. Additionally, the notional proportion, $\hat{P}^{+|k=3,r,i}_0$. Table restricted to $r \in [7, 8, 9, 10, 11]$ and $r - 3 \leq i \leq r - 1$. BY-corrected $p$-values shown in parentheses.}

\centering
\begin{tabular}{rrcrrrr}
\toprule
$r$ & $i$ & \text{Win series?} & $r^{+|k=3,r,i}$ & $N^{k=3,r,i}$ & $P^{+|k=3,r,i}$ \% & $\hat{P}^{+|k=3,r,i}_0$ \% \\ 
\midrule

7 & 5 & yes & 36 & 54 & 66.7 & 66.7 (1.00) \\ 
7 & 6 & yes & 34 & 54 & 63.0 & 50.0 (1.00) \\ 
8 & 6 & yes & 20 & 41 & 48.8 & 50.0 (1.00) \\ 
8 & 7 & yes & 24 & 46 & 52.2 & 50.0 (1.00) \\ 
9 & 7 & yes & 15 & 33 & 45.5 & 40.0 (1.00) \\ 
9 & 8 & yes & 18 & 37 & 48.6 & 50.0 (1.00) \\ 
10 & 8 & no & 1 & 13 & 7.7 & 33.3 (1.00) \\ 
10 & 8 & yes & 9 & 24 & 37.5 & 33.3 (1.00) \\ 
10 & 9 & no & 2 & 12 & 16.7 & 50.0 (1.00) \\ 
10 & 9 & yes & 4 & 20 & 20.0 & 50.0 (1.00) \\ 
11 & 9 & no & 5 & 12 & 41.7 & 28.6 (1.00) \\ 
11 & 9 & yes & 10 & 20 & 50.0 & 28.6 (1.00) \\ 
11 & 10 & no & 5 & 14 & 35.7 & 50.0 (1.00) \\ 
11 & 10 & yes & 5 & 22 & 22.7 & 50.0 (1.00) \\

@@

In Table \ref{tbl:ww-sequences} the observed and expected count of
sequences---\(P^{\zeta}_j\) and \(\hat{P}^{\zeta}_j\) respectively---are
shown for sequences for which we can reject the Wald-Wolfowitz null
hypothesis\footnote{Interestingly, we identify the same non-random
  sequences when using either \(\phi = 0.575\) or \(\phi = 0.5\).} at a
confidence level of \(\alpha = 0.05\), for
\(r \in [6, 7, 8, 9, 10]\)\footnote{Since the normal method is invalid
  for sequences with only successes or failures, we use the exact method
  to evaluate (and reject) the null hypothesis for the Wald-Wolfowitz
  runs test for \(r = 6\).}. In addition to the 14 sequences shown,
there are 12 additional sequences for \(r = 11\), not shown for
brevity.\footnote{If we had used the exact method for detecting
  non-random sequences with the Wald-Wolfowitz runs test, we would have
  identified 20 sequences total.}

@@

\caption{Sequences ($\zeta_j$) for which we can reject Wald-Wolfowitz null hypothesis for $r \in [6, 7, 8, 9, 10]$, where round wins and losses are denoted with $+$ and $-$ respectively. Additionally, the observed count, $N^{\zeta}_j$, and proportion, $P^{\zeta}_j$, of all possible sequences, as well as the expected count, $\hat{N}^{\zeta}_j$, and proportion, $\hat{P}^{\zeta}_j$. $\hat{N}^{\zeta}_j$ is scaled to the observed number of series played, hence its non-integer value. BY-corrected equal proportions test $p$-values in parentheses.}

\centering
\begin{tabular}{rlrrrr}
\toprule
$r$ & $\zeta_j$ & $N^{\zeta}_j$ & $P^{\zeta}_j$ \% & $\hat{N}^{\zeta}_j$ & $\hat{P}^{\zeta}_j$ \% \\ 
\midrule

6 & + + + + + + & 40 & 4.69 & 386 & 3.86 (0.62) \\ 
8 & - - + + + + + + & 4 & 0.47 & 90 & 0.90 (0.23) \\ 
9 & - - - + + + + + + & 4 & 0.47 & 44 & 0.44 (1.00) \\ 
10 & - - - - + + + + + + & 2 & 0.23 & 21 & 0.21 (1.00) \\ 
10 & + - - - - + + + + + & 0 & 0.00 & 24 & 0.24 (0.23) \\ 
10 & + - + - + - + - + + & 3 & 0.35 & 13 & 0.13 (0.62) \\ 
10 & + - + - + - + + - + & 3 & 0.35 & 20 & 0.20 (1.00) \\ 
10 & + - + - + + - + - + & 3 & 0.35 & 24 & 0.24 (1.00) \\ 
10 & + - + + - + - + - + & 0 & 0.00 & 20 & 0.20 (0.23) \\ 
10 & + + - - - - + + + + & 1 & 0.12 & 14 & 0.14 (1.00) \\ 
10 & + + - + - + - + - + & 0 & 0.00 & 13 & 0.13 (0.48) \\ 
10 & + + + - - - - + + + & 2 & 0.23 & 24 & 0.24 (1.00) \\ 
10 & + + + + - - - - + + & 1 & 0.12 & 28 & 0.28 (0.61) \\ 
10 & + + + + + - - - - + & 2 & 0.23 & 12 & 0.12 (1.00) \\
